---
title: "Use Actual Stock Assessment"
author: "Chengxue Li"
date: "Last update: `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header.html
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5
)
```

This vignette demonstrates how to use an **actual stock assessment model**
output to develop an operating model (OM) for `whamMSE`, while keeping the
historical population dynamics fixed to the fitted stock trajectory.

In many MSE examples, process random effects (for example, recruitment
deviations) are simulated for the entire time series. That means the
historical population trajectory can vary from one random seed to another.
However, when an OM is built from a real stock assessment, users may prefer
the historical period to reflect the fitted trajectory of the real stock as
closely as possible. This can make the evaluation of management procedures
more meaningful for management applications.

The key idea in this vignette is:

1. **Fit a stock assessment model to real data**
   - Here we use Georges Bank yellowtail flounder data.

2. **Extract estimated parameters and historical process deviations**
   - These are used to construct an OM for MSE.

3. **Fix the historical period**
   - The historical trajectory is held constant across seeds, so the
     underlying population dynamics in the historical period do not change.

4. **Simulate only the feedback period**
   - Random deviations are still drawn during the feedback years, so future
     trajectories can differ across seeds.

At present, this feature is available only in the **lab branch** of
`whamMSE`. In addition, this workflow has only been tested for cases where
**recruitment is the only process treated as a random effect in the OM**.
Users should therefore use an OM with **recruitment-only random effects**
(i.e., no NAA random effects in the OM). Other process variation, such as
natural mortality, may still be included as fixed process structure. In the
EM, users may still include any process as random effects if desired.

This vignette provides a full end-to-end example:

- fit the historical stock assessment,
- build the projection OM,
- fix the historical period,
- simulate the future,
- run MSE.

This is an illustrative example using **Georges Bank yellowtail flounder**.

### 1. Load `WHAM` and `whamMSE`

```{r, eval=TRUE}
# Install if needed.
# This fixed-history workflow is currently available in the lab branch.
# remotes::install_github("lichengxue/whamMSE@lab", upgrade = "never")

library(wham)
library(whamMSE)
library(dplyr)
library(ggplot2)
library(here)

main.dir <- here::here()
```

### 2. Read stock assessment data and define global settings

We use Georges Bank yellowtail flounder ASAP input data and set up a short MSE
feedback period.

```{r, eval=TRUE}
gb_dat <- read_asap3_dat("Data/GBK.DAT")

input_hist <- prepare_wham_input(gb_dat)
input      <- prepare_wham_input(gb_dat)

n_stocks  <- 1L
n_regions <- 1L
n_ages    <- 6L
n_indices <- 3L
n_fleets  <- 1L
n_seasons <- 1L

year_start <- 1973L
year_end   <- 2022L
MSE_years  <- 10L

n_hist_years <- year_end - year_start + 1L
n_proj_years <- n_hist_years + MSE_years

base_years      <- year_start:year_end
assess_interval <- 2L

stopifnot(MSE_years %% assess_interval == 0L)

assess_years <- seq(
  tail(base_years, 1),
  year_end + MSE_years - assess_interval,
  by = assess_interval
)

# Historical period is fixed.
# Only years from first_free_year onward are allowed to vary.
first_free_year <- 51L
```

### 3. Fit the historical stock assessment model

We first fit a historical OM to the real stock assessment data. In this
example, recruitment is the only process treated as random effects in the
historical fit.

#### 3.1 Biological inputs for the historical period

```{r, eval=TRUE}
user_maturity_hist <- array(NA_real_, dim = c(n_stocks, n_hist_years, n_ages))
user_maturity_hist[, 1:n_hist_years, ] <- input$data$mature

user_waa_hist <- list()
user_waa_hist$waa <- array(NA_real_, dim = c(5, n_hist_years, n_ages))
user_waa_hist$waa[, 1:n_hist_years, ] <- input$data$waa
user_waa_hist$waa_pointer_fleets   <- input$data$waa_pointer_fleets
user_waa_hist$waa_pointer_indices  <- input$data$waa_pointer_indices
user_waa_hist$waa_pointer_totcatch <- input$data$waa_pointer_ssb
user_waa_hist$waa_pointer_ssb      <- input$data$waa_pointer_ssb
user_waa_hist$waa_pointer_M        <- input$data$waa_pointer_M

fracyr_spawn <- gb_dat[[1]]$dat$fracyr_spawn
```

#### 3.2 Observation settings for catch and indices

```{r, eval=TRUE}
catch_info_hist <- list(
  catch_cv      = 0.05,
  catch_Neff    = 50,
  use_agg_catch = 1,
  use_catch_paa = 1
)

index_info_hist <- list(
  index_cv        = rep(0.5, n_indices),
  index_Neff      = rep(25, n_indices),
  fracyr_indices  = c(0.25, 0.75, 0.25),
  q               = rep(0.2, n_indices),
  use_indices     = rep(1, n_indices),
  use_index_paa   = rep(1, n_indices),
  units_indices   = rep(2, n_indices),
  units_index_paa = rep(2, n_indices)
)
```

#### 3.3 Build the historical model scaffold

```{r, eval=TRUE}
info_hist <- whamMSE::generate_basic_info(
  n_stocks = n_stocks,
  n_regions = n_regions,
  n_indices = n_indices,
  n_fleets = n_fleets,
  n_seasons = n_seasons,
  base.years = year_start:year_end,
  n_feedback_years = 0L,
  n_ages = n_ages,
  catch_info = catch_info_hist,
  index_info = index_info_hist,
  user_waa = user_waa_hist,
  user_maturity = user_maturity_hist,
  fracyr_spawn = fracyr_spawn
)

basic_info_hist     <- info_hist$basic_info
catch_info_use_hist <- info_hist$catch_info
index_info_use_hist <- info_hist$index_info
F_info_hist         <- info_hist$F

catch_info_use_hist$agg_catch[1:n_hist_years, ]     <- input$data$agg_catch
catch_info_use_hist$catch_paa[, 1:n_hist_years, ]   <- input$data$catch_paa
index_info_use_hist$agg_indices[1:n_hist_years, ]   <- input$data$agg_indices
index_info_use_hist$index_paa[, 1:n_hist_years, ]   <- input$data$index_paa
index_info_use_hist$use_indices[1:n_hist_years, ]   <- input$data$use_indices
index_info_use_hist$use_index_paa[1:n_hist_years, ] <- input$data$use_index_paa
```

#### 3.4 Selectivity, natural mortality, and recruitment process

```{r, eval=TRUE}
sel_hist <- list(
  model = c("age-specific", "logistic", "logistic", "logistic"),
  re    = c("ar1_y", "none", "none", "none"),
  initial_pars = list(
    c(0.1, 0.25, 0.5, 1, 1, 1),
    c(2, 0.3),
    c(2, 0.3),
    c(2, 0.3)
  ),
  fix_pars = list(c(4:6), NULL, NULL, NULL)
)

M_hist <- list(
  model = "constant",
  initial_means = array(
    c(0.57, 0.33, 0.26, 0.23, 0.22, 0.22),
    dim = c(n_stocks, n_regions, n_ages)
  )
)

sigma_vals_hist <- array(0, dim = c(n_stocks, n_regions, n_ages))
sigma_vals_hist[, , 1] <- 0.5

NAA_re_hist <- list(
  N1_model      = rep("equilibrium", n_stocks),
  sigma         = rep("rec", n_stocks),
  cor           = rep("iid", n_stocks),
  recruit_model = 3,
  sigma_vals    = sigma_vals_hist
)
```

#### 3.5 Prepare and fit the historical model

```{r, eval=TRUE}
input_histfit <- prepare_wham_input(
  basic_info  = basic_info_hist,
  selectivity = sel_hist,
  M           = M_hist,
  NAA_re      = NAA_re_hist,
  catch_info  = catch_info_use_hist,
  index_info  = index_info_use_hist,
  F           = F_info_hist,
  age_comp    = "logistic-normal-pool0"
)

input_histfit <- update_waa(input_histfit, waa_info = info_hist$par_inputs$user_waa)

input_e <- prepare_wham_input(gb_dat)
input_histfit$data$agg_index_sigma[1:n_hist_years, ] <- input_e$data$agg_index_sigma
input_histfit$data$use_indices[1:n_hist_years, ]     <- input_e$data$use_indices
input_histfit$data$use_index_paa[1:n_hist_years, ]   <- input_e$data$use_index_paa
```

```{r, eval=TRUE}
generate_remove_years <- function(mat) {
  lz <- lapply(seq_len(ncol(mat)), function(j) which(mat[, j] == 0))
  if (all(lengths(lz) == 0)) return(NULL)
  max_rows <- max(sapply(lz, length))
  out <- matrix(NA_integer_, nrow = max_rows, ncol = length(lz))
  for (j in seq_along(lz)) {
    if (length(lz[[j]])) out[seq_along(lz[[j]]), j] <- lz[[j]]
  }
  out
}

remove_agg_years1 <- generate_remove_years(input_e$data$use_indices)
remove_paa_years1 <- generate_remove_years(input_e$data$use_index_paa)

input_histfit <- update_input_index_info(
  input_histfit,
  agg_index_sigma = input_histfit$data$agg_index_sigma,
  index_Neff      = input_histfit$data$index_Neff,
  remove_agg = TRUE,
  remove_agg_pointer = 1:3,
  remove_agg_years = remove_agg_years1,
  remove_paa = TRUE,
  remove_paa_pointer = 1:3,
  remove_paa_years = remove_paa_years1
)
```

```{r, eval=TRUE}
if (!file.exists("Data/OM.RDS")) {
  OM <- fit_wham(
    input_histfit,
    do.fit = TRUE,
    do.brps = TRUE,
    do.retro = FALSE,
    do.osa = FALSE,
    MakeADFun.silent = TRUE
  )
  saveRDS(OM, "Data/OM.RDS")
} else {
  OM <- readRDS("Data/OM.RDS")
}

check_convergence(OM)
```

### 4. Build the projection operating model

Next, we extend the model from the historical period into the feedback period.
The historical years will later be fixed to the fitted stock trajectory, while
the feedback years will be allowed to vary by seed.

#### 4.1 Biological inputs over the full modeled period

```{r, eval=TRUE}
user_maturity <- array(NA_real_, dim = c(n_stocks, n_proj_years, n_ages))
user_maturity[, 1:n_hist_years, ] <- OM$input$data$mature[, 1:n_hist_years, ]
for (iy in (n_hist_years + 1L):n_proj_years) {
  user_maturity[, iy, ] <- OM$input$data$mature[, n_hist_years, , drop = FALSE]
}

user_waa <- list()
user_waa$waa <- array(NA_real_, dim = c(5, n_proj_years, n_ages))
user_waa$waa[, 1:n_hist_years, ] <- OM$input$data$waa[, 1:n_hist_years, ]
for (iy in (n_hist_years + 1L):n_proj_years) {
  user_waa$waa[, iy, ] <- OM$input$data$waa[, n_hist_years, , drop = FALSE]
}

user_waa$waa_pointer_fleets   <- OM$input$data$waa_pointer_fleets
user_waa$waa_pointer_indices  <- OM$input$data$waa_pointer_indices
user_waa$waa_pointer_totcatch <- OM$input$data$waa_pointer_totcatch
user_waa$waa_pointer_ssb      <- OM$input$data$waa_pointer_ssb
user_waa$waa_pointer_M        <- OM$input$data$waa_pointer_M
```

#### 4.2 Catch and index settings for the OM

```{r, eval=TRUE}
catch_info <- list(
  catch_cv      = 0.05,
  catch_Neff    = 50,
  use_agg_catch = 1,
  use_catch_paa = 1
)

index_info <- list(
  index_cv        = rep(0.5, n_indices),
  index_Neff      = rep(25, n_indices),
  fracyr_indices  = c(0.25, 0.75, 0.25),
  q               = c(2.110e-4, 2.247e-4, 2.683e-4),
  use_indices     = rep(1, n_indices),
  use_index_paa   = rep(1, n_indices),
  units_indices   = rep(2, n_indices),
  units_index_paa = rep(2, n_indices)
)
```

#### 4.3 Build the projection scaffold

```{r, eval=TRUE}
info <- whamMSE::generate_basic_info(
  n_stocks = n_stocks,
  n_regions = n_regions,
  n_indices = n_indices,
  n_fleets = n_fleets,
  n_seasons = n_seasons,
  base.years = year_start:year_end,
  n_feedback_years = MSE_years,
  n_ages = n_ages,
  catch_info = catch_info,
  index_info = index_info,
  user_waa = user_waa,
  user_maturity = user_maturity,
  fracyr_spawn = fracyr_spawn
)

basic_info     <- info$basic_info
catch_info_use <- info$catch_info
index_info_use <- info$index_info
```

#### 4.4 Initialize fishing mortality

```{r, eval=TRUE}
F_info <- info$F
F_info$F[1:n_hist_years, ] <- OM$rep$Fbar[1:n_hist_years, 1, drop = FALSE]
```

#### 4.5 Define selectivity, M, and process settings for the OM

Although the historical fit used recruitment-only random effects, the
projection OM uses `rec+1` so that the historical stock trajectory can be
reconstructed more closely when fixed values are inserted.

```{r, eval=TRUE}
sel3 <- list(
  model = c("age-specific", "logistic", "logistic", "logistic"),
  re    = c("ar1_y", "none", "none", "none"),
  initial_pars = list(
    c(0.05, 0.3, 0.7, 0.875, 0.9, 1),
    c(2.305, 0.327),
    c(1.611, 0.483),
    c(2.135, 0.217)
  ),
  fix_pars = list(c(4:6), NULL, NULL, NULL)
)

M <- list(
  model = "constant",
  initial_means = array(
    c(0.57, 0.33, 0.26, 0.23, 0.22, 0.22),
    dim = c(n_stocks, n_regions, n_ages)
  )
)

sigma_vals <- array(0, dim = c(n_stocks, n_regions, n_ages))
sigma_vals[, , 1] <- exp(OM$parList$log_NAA_sigma[, , 1])
sigma_vals[, , 2:n_ages] <- exp(OM$parList$log_NAA_sigma[, , 2:n_ages])

recruit_pars_use <- exp(as.numeric(OM$parList$mean_rec_pars[1, ]))

NAA_re <- list(
  N1_model      = rep("age-specific-fe", n_stocks),
  sigma         = rep("rec+1", n_stocks),
  cor           = rep("iid", n_stocks),
  recruit_model = 3,
  recruit_pars  = list(recruit_pars_use),
  sigma_vals    = sigma_vals
)
```

#### 4.6 Build the unfitted projection OM input

```{r, eval=TRUE}
om_input_P <- wham::prepare_wham_input(
  basic_info  = basic_info,
  selectivity = sel3,
  M           = M,
  NAA_re      = NAA_re,
  catch_info  = catch_info_use,
  index_info  = index_info_use,
  F           = F_info,
  age_comp    = "logistic-normal-pool0"
)

waa_info <- info$par_inputs$user_waa
om_input_P <- whamMSE::update_waa(om_input_P, waa_info = waa_info)

om_input_P$par$catch_paa_pars <- OM$parList$catch_paa_pars
om_input_P$par$index_paa_pars <- OM$parList$index_paa_pars
om_input_P$par$sel_repars     <- OM$parList$sel_repars

for (i in 1:n_regions) {
  om_input_P$par$log_N1[i, i, ] <- log(OM$rep$N1)[i, i, ]
}

om_input_P$data$agg_index_sigma[1:n_hist_years, ] <- input_hist$data$agg_index_sigma
om_input_P$data$use_indices[1:n_hist_years, ]     <- input_hist$data$use_indices
om_input_P$data$use_index_paa[1:n_hist_years, ]   <- input_hist$data$use_index_paa

for (iy in (n_hist_years + 1L):n_proj_years) {
  om_input_P$data$agg_index_sigma[iy, ] <- input_hist$data$agg_index_sigma[n_hist_years, ]
}
```

```{r, eval=TRUE}
loc_zeros_agg <- lapply(seq_len(ncol(input_hist$data$use_indices)), function(j) {
  which(input_hist$data$use_indices[, j] == 0)
})

loc_zeros_paa <- lapply(seq_len(ncol(input_hist$data$use_index_paa)), function(j) {
  which(input_hist$data$use_index_paa[, j] == 0)
})

max_agg <- max(sapply(loc_zeros_agg, length))
max_paa <- max(sapply(loc_zeros_paa, length))

remove_agg_years1 <- if (max_agg > 0) {
  out <- matrix(NA_integer_, nrow = max_agg, ncol = length(loc_zeros_agg))
  for (j in seq_along(loc_zeros_agg)) {
    if (length(loc_zeros_agg[[j]])) {
      out[seq_along(loc_zeros_agg[[j]]), j] <- loc_zeros_agg[[j]]
    }
  }
  out
} else NULL

remove_paa_years1 <- if (max_paa > 0) {
  out <- matrix(NA_integer_, nrow = max_paa, ncol = length(loc_zeros_paa))
  for (j in seq_along(loc_zeros_paa)) {
    if (length(loc_zeros_paa[[j]])) {
      out[seq_along(loc_zeros_paa[[j]]), j] <- loc_zeros_paa[[j]]
    }
  }
  out
} else NULL

om_input_P <- whamMSE::update_input_index_info(
  om_input_P,
  agg_index_sigma = om_input_P$data$agg_index_sigma,
  index_Neff      = om_input_P$data$index_Neff,
  remove_agg = TRUE,
  remove_agg_pointer = 1:3,
  remove_agg_years = remove_agg_years1,
  remove_paa = TRUE,
  remove_paa_pointer = 1:3,
  remove_paa_years = remove_paa_years1
)
```

### 5. Compile the unfitted projection OM

```{r, eval=TRUE}
om_input_P$data$do_simulate_N_re <- 1L

unfitted_om_P <- fit_wham(
  om_input_P,
  do.fit = FALSE,
  do.brps = FALSE,
  MakeADFun.silent = TRUE
)
```

### 6. Fix the historical period and free only the feedback period

This is the key step. We overwrite the historical `log_NAA` values using the
fitted historical model so that the historical population trajectory remains
fixed across seeds. Only the future block is left free and can be simulated.

```{r, eval=TRUE}
unfitted_om_P$input$random <- "log_NAA"
unfitted_om_P$input$data$do_simulate_N_re <- 1L

unfitted_om_P$input$par$log_N1 <- log(OM$rep$N1)
if (!is.null(unfitted_om_P$parList$log_N1)) {
  unfitted_om_P$parList$log_N1 <- log(OM$rep$N1)
}

hist_idx <- 1:(first_free_year - 2L)
unfitted_om_P$input$par$log_NAA[1, 1, hist_idx, ] <-
  log(OM$rep$NAA[1, 1, hist_idx + 1L, ])

if (!is.null(unfitted_om_P$parList$log_NAA)) {
  unfitted_om_P$parList$log_NAA[1, 1, hist_idx, ] <-
    log(OM$rep$NAA[1, 1, hist_idx + 1L, ])
}
```

```{r, eval=TRUE}
logNAA_dim <- dim(unfitted_om_P$input$par$log_NAA)
future_idx_start <- first_free_year - 1L
future_idx <- future_idx_start:logNAA_dim[3]

m <- array(NA_integer_, dim = logNAA_dim)
m[1, 1, future_idx, ] <- 1L
m[!is.na(m)] <- seq_len(sum(!is.na(m)))

unfitted_om_P$input$map$log_NAA <- factor(m)
if (!is.null(unfitted_om_P$env$map)) {
  unfitted_om_P$env$map$log_NAA <- unfitted_om_P$input$map$log_NAA
}
```

```{r, eval=TRUE}
last_fixed_idx <- first_free_year - 2L
if (last_fixed_idx >= 1L) {
  unfitted_om_P$input$par$log_NAA[1, 1, future_idx, ] <-
    array(
      unfitted_om_P$input$par$log_NAA[1, 1, last_fixed_idx, ],
      dim = c(1, 1, length(future_idx), logNAA_dim[4])
    )

  if (!is.null(unfitted_om_P$parList$log_NAA)) {
    unfitted_om_P$parList$log_NAA[1, 1, future_idx, ] <-
      array(
        unfitted_om_P$parList$log_NAA[1, 1, last_fixed_idx, ],
        dim = c(1, 1, length(future_idx), logNAA_dim[4])
      )
  }
}

unfitted_om_P$parList$logit_selpars <- OM$parList$logit_selpars
unfitted_om_P$parList$selpars_re    <- OM$parList$selpars_re
```

```{r, eval=TRUE}
unfitted_om_P <- fit_wham(
  unfitted_om_P$input,
  do.fit = FALSE,
  do.brps = FALSE,
  MakeADFun.silent = TRUE
)
```

### 7. Simulate only the future block

Now we simulate the OM while keeping the historical period fixed. The future
trajectory depends on the seed, but the historical trajectory remains
unchanged.

```{r, eval=TRUE}
seed <- 12345L

om_future <- update_om_fn(
  unfitted_om_P,
  seed = seed,
  random = "log_NAA",
  process_fix = 1,
  first_free_year = first_free_year
)
```

### 8. Diagnostics

The historical trajectory should closely match the fitted OM, while the future
trajectory is free to vary.

```{r, eval=TRUE}
plot(
  om_future$rep$SSB,
  type = "l",
  main = "SSB with fixed history and simulated feedback period",
  xlab = "Year index",
  ylab = "SSB"
)
lines(1:n_hist_years, OM$rep$SSB[1:n_hist_years], col = "red", lwd = 2)

legend(
  "topright",
  legend = c("Projected OM", "Fitted historical OM"),
  col = c("black", "red"),
  lwd = c(1, 2),
  bty = "n"
)
```

```{r, eval=TRUE}
s <- 1
r <- 1
hist_idx <- 1:(n_hist_years - 1L)

par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))

plot(
  hist_idx + 1,
  exp(om_future$input$par$log_NAA[s, r, hist_idx, 1]),
  type = "l",
  xlab = "Model year",
  ylab = "NAA age 1",
  main = "Historical check: input log_NAA vs fitted OM rep$NAA"
)
lines(
  hist_idx + 1,
  OM$rep$NAA[s, r, hist_idx + 1L, 1],
  col = "red"
)

plot(
  om_future$rep$pred_NAA[s, r, , 6],
  type = "l",
  main = "pred_NAA age 6",
  xlab = "Year index",
  ylab = "pred_NAA age 6"
)
abline(v = first_free_year, lty = 2, col = "red")
```

```{r, eval=TRUE}
plot(om_future$rep$NAA_devs[, , , 1], type = "l", main = "NAA_devs age 1")
plot(om_future$rep$pred_NAA[, , , 1], type = "l", main = "pred_NAA age 1")
```

```{r, eval=TRUE}
hist_diff <- om_future$rep$NAA[, , 1:n_hist_years, ] -
  OM$rep$NAA[, , 1:n_hist_years, ]
print(hist_diff)

plot(
  om_future$rep$SSB[1:n_hist_years],
  type = "l",
  main = "Historical SSB comparison",
  xlab = "Year index",
  ylab = "SSB"
)
lines(OM$rep$SSB[1:n_hist_years], col = "red")
```

### 9. Run MSE

We now run MSE using the projected OM with fixed historical dynamics. In this
example, we compare two harvest control rule settings: 75% FXSPR and 100%
FXSPR.

#### 9.1 Define assessment years and EM settings

```{r, eval=TRUE}
assess_interval <- 2L
base_years      <- year_start:year_end
terminal_year   <- tail(base_years, 1)

assess_years <- seq(
  terminal_year,
  year_end + MSE_years - assess_interval,
  by = assess_interval
)

NAA_re_em <- NAA_re
NAA_re_em$N1_model <- "equilibrium"
NAA_re_em$sigma    <- "rec"
NAA_re_em$cor      <- "iid"

update_catch_info_list <- list(
  agg_catch_sigma = om_input_P$data$agg_catch_sigma,
  catch_Neff      = om_input_P$data$catch_Neff
)

update_index_info_list <- list(
  agg_index_sigma = om_input_P$data$agg_index_sigma,
  index_Neff      = om_input_P$data$index_Neff,
  remove_agg = TRUE,
  remove_agg_pointer = 1:3,
  remove_agg_years = remove_agg_years1,
  remove_paa = TRUE,
  remove_paa_pointer = 1:3,
  remove_paa_years = remove_paa_years1
)
```

#### 9.2 MSE run with 75% FXSPR

```{r, eval=TRUE}
hcr1 <- list(
  hcr.type = 1,
  hcr.opts = list(use_FXSPR = TRUE, percentFXSPR = 75)
)

if (!file.exists("Results/mod1.RDS")) {
  mod1 <- loop_through_fn(
    om = om_future,
    em_info = info,
    random = "log_NAA",
    M_em = M,
    sel_em = sel3,
    NAA_re_em = NAA_re_em,
    age_comp_em = "logistic-normal-pool0",
    em.opt = list(
      separate.em = FALSE,
      separate.em.type = 1,
      do.move = FALSE,
      est.move = FALSE
      ),
    update_index_info = update_index_info_list,
    update_catch_info = update_catch_info_list,
    assess_years = assess_years,
    assess_interval = assess_interval,
    base_years = base_years,
    year.use = 50,
    add.years = TRUE,
    seed = 12345L,
    hcr = hcr1,
    save.sdrep = FALSE,
    save.last.em = TRUE,
    FXSPR_init = 0.5,
    process_fix = 1,
    first_free_year = 51
    ) 
  saveRDS(mod1, "Results/mod1.RDS")
  } else {
    mod1 <- readRDS("Results/mod1.RDS")
  }
```

#### 9.3 MSE run with 100% FXSPR

```{r, eval=TRUE}
hcr2 <- list(
  hcr.type = 1,
  hcr.opts = list(use_FXSPR = TRUE, percentFXSPR = 100)
)

if (!file.exists("Results/mod2.RDS")) {
  mod2 <- loop_through_fn(
    om = om_future,
    em_info = info,
    random = "log_NAA",
    M_em = M,
    sel_em = sel3,
    NAA_re_em = NAA_re_em,
    age_comp_em = "logistic-normal-pool0",
    em.opt = list(
      separate.em = FALSE,
      separate.em.type = 1,
      do.move = FALSE,
      est.move = FALSE
      ),
    update_index_info = update_index_info_list,
    update_catch_info = update_catch_info_list,
    assess_years = assess_years,
    assess_interval = assess_interval,
    base_years = base_years,
    year.use = 50,
    add.years = TRUE,
    seed = 12345L,
    hcr = hcr2,
    save.sdrep = FALSE,
    save.last.em = TRUE,
    FXSPR_init = 0.5,
    process_fix = 1,
    first_free_year = 51
    )
  saveRDS(mod2, "Results/mod2.RDS")
  } else {
    mod2 <- readRDS("Results/mod2.RDS")
  }
```

### 10. Plot MSE results

Below we compare the realized OM trajectories under the two HCR settings.

```{r, eval=TRUE}
plot(
  mod1$om$rep$SSB,
  type = "l",
  lwd = 2,
  col = "blue",
  xlab = "Year index",
  ylab = "SSB",
  main = "MSE comparison: SSB under two HCR settings"
)

lines(
  mod2$om$rep$SSB,
  col = "red",
  lwd = 2,
  lty = 2
)

abline(v = first_free_year, lty = 3)

legend(
  "topright",
  legend = c("75% FXSPR", "100% FXSPR", "Start of feedback period"),
  col = c("blue", "red", "black"),
  lty = c(1, 2, 3),
  lwd = c(2, 2, 1),
  bty = "n"
)
```

### 11. Summary

This vignette showed how to build an MSE operating model directly from a fitted
stock assessment model while keeping the historical population trajectory fixed.

The main advantage of this workflow is that the historical period remains
anchored to the fitted trajectory of the real stock, rather than changing
across seeds because of newly simulated historical process deviations. This can
be especially useful when the goal is to evaluate management procedures under a
historical stock trajectory that is already scientifically accepted or
management-relevant.

In this example:

- the historical stock assessment was fit to real Georges Bank yellowtail
  flounder data,
- estimated parameters and historical trajectory were used to construct the OM,
- the historical period was fixed,
- only the feedback period was simulated,
- MSE was run under two HCR settings.

At present, users should apply this workflow only to OMs where **recruitment is
the only random process tested in the historical fixing workflow**.
