---
title: "Parallel Computing"
author: "Chengxue Li"
date: "Last update: `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header.html
editor_options: 
  markdown: 
    wrap: 72
---

This vignette includes an example of how to use parallel computing to conduct management strategy evaluation.

### 1. Load "WHAM" and "whamMSE":

```{r, eval=FALSE}
library(wham)
library(whamMSE)

main.dir = here::here()
```

### 2. Generate basic information

```{r, eval=FALSE}
year_start  <- 2003  # starting year in the burn-in period
year_end    <- 2022  # end year in the burn-in period
MSE_years   <- 3     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation 

info <- generate_basic_info(n_stocks   = 2, 
                            n_regions  = 2, 
                            n_indices  = 2, 
                            n_fleets   = 2, 
                            n_seasons  = 4, 
                            base.years = year_start:year_end, 
                            n_feedback_years = MSE_years, 
                            life_history  = "medium", 
                            n_ages        = 12, 
                            Fbar_ages     = 12, 
                            recruit_model = 2, 
                            F_info     = list(F.year1 = 0.2, Fhist = "updown",
                                              Fmax = 1, Fmin = 0.2,change_time = 0.5),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 200), 
                            index_info = list(index_cv = 0.1, index_Neff = 200, 
                                              fracyr_indices = 0.5, q = 0.2), 
                            fracyr_spawn = 0.5, 
                            bias.correct.process     = FALSE, 
                            bias.correct.observation = FALSE, 
                            bias.correct.BRPs        = FALSE, 
                            mig_type = 0) 

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

### 3. Specify movement type and movement rate

```{r, eval=FALSE}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2) # "bidirectional" movement

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = c(0.3,0.1), move.re = "constant")

```

### 4. Configure selecitvity and natural mortality

```{r, eval=FALSE}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
```

### 5. Configure numbers-at-age (NAA)

```{r, eval=FALSE}
sigma        <- "rec+1"
re_cor       <- "iid"
ini.opt      <- "equilibrium" # option <- c("age-specific-fe", "equilibrium")
Rec_sig      <- 0.5 # (sigma for recruitment)
NAA_sig      <- 0.5 # (sigma for NAA)

sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages
sigma_vals[,,1] = Rec_sig

# Set initial NAA for each stock
log_N1    <- rep(10,n_stocks)
log_N1[1] <- log(exp(10)*2) # N1_stock1 is 2 times higher than N1_stock2
N1_pars   <- generate_ini_N1(log_N1,basic_info,ini.opt)

# Set mean recruitment para. for each stock
mean_rec_par <- list()
for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i])

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 2,  # rec random around the mean
               recruit_pars = mean_rec_par, 
               sigma_vals = sigma_vals,  
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where) # NAA_where must be specified in basic_info!
```

### 6. Generate wham input

Here we use `prepare_wham_input()` function to generate a wham input
using the basic information we set above:

```{r, eval=FALSE}
input <- prepare_wham_input(basic_info = basic_info, selectivity = sel, M = M, NAA_re = NAA_re, move = move, catch_info = catch_info, index_info = index_info, F = F_info)
```

### 7. Generate operating model

```{r, eval=FALSE}
random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

### 8. Specify assessment interval and assessment year in the feedback loop

```{r, eval=FALSE}
assess.interval <- 3 # Note: assessment interval is 3 years, given the feedback period is 3 years, there will be only 1 assessment
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs
```

### 9. Create a folder to save the results
```{r, eval=FALSE}
sub.dir = "Results"
dir.create(file.path(getwd(), sub.dir), recursive = TRUE)

library(doParallel)
library(foreach)

detectCores() # check how many cores available
```

### 10. EM1
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
  
  n_stocks = n_regions = n_fleets = n_indices = 1
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  mod <- loop_through_fn(om = om_with_data, 
                         em_info = info,
                         random = random,
                         M_em = M_em, 
                         sel_em = sel_em,
                         NAA_re_em = NAA_re_em, 
                         move_em = NULL,
                         age_comp_em = "multinomial",
                         em.opt = list(separate.em = TRUE, separate.em.type = 1, 
                                       do.move = FALSE, est.move = FALSE),
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 20, 
                         hcr.type = 1,
                         seed = 123+i,
                         save.sdrep = FALSE)
  
    saveRDS(mod, file.path(sub.dir,sprintf("Mod1_%03d.RDS",i)))

}

stopCluster(cluster)
```

### 11. EM2
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
  
  n_stocks = n_regions = 1
  n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  mod <- loop_through_fn(om = om_with_data, 
                         em_info = info,
                         random = random,
                         M_em = M_em, 
                         sel_em = sel_em, 
                         NAA_re_em = NAA_re_em, 
                         move_em = NULL,
                         age_comp_em = "multinomial",
                         em.opt = list(separate.em = TRUE, separate.em.type = 2, 
                                       do.move = FALSE, est.move = FALSE),
                         assess_years = assess.years,
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 20,
                         hcr.type = 1,
                         seed = 123+i,
                         save.sdrep = FALSE)
  
  saveRDS(mod, file.path(sub.dir,sprintf("Mod2_%03d.RDS",i)))
  
}

stopCluster(cluster)
```

### 12. EM3
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
  
  n_stocks = n_regions = n_fleets = n_indices = 1
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
                 initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  mod <- loop_through_fn(om = om_with_data, 
                         em_info = info,
                         random = random,
                         M_em = M_em, 
                         sel_em = sel_em, 
                         NAA_re_em = NAA_re_em, 
                         move_em = NULL,
                         age_comp_em = "multinomial", 
                         em.opt = list(separate.em = TRUE, separate.em.type = 3, 
                                       do.move = FALSE, est.move = FALSE),
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 20, 
                         hcr.type = 1,
                         seed = 123+i,
                         save.sdrep = FALSE)
  saveRDS(mod, file.path(sub.dir,sprintf("Mod3_%03d.RDS",i)))
}

stopCluster(cluster)
```

### 13. EM4
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
  
  n_stocks = n_regions = n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

  NAA_re_em <- list(N1_model=rep("equilibrium",n_stocks),
                  sigma=rep("rec+1",n_stocks),
                  cor=rep("iid",n_stocks),
                  recruit_model = 2)

  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  mod <- loop_through_fn(om = om_with_data, 
                         em_info = info,
                         random = random,
                         M_em = M_em, 
                         sel_em = sel_em, 
                         NAA_re_em = NAA_re_em, 
                         move_em = NULL,
                         age_comp_em = "multinomial",
                         em.opt = list(separate.em = FALSE, separate.em.type = NULL, 
                                       do.move = FALSE,  est.move = FALSE),
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 20,
                         hcr.type = 1,
                         seed = 123+i,
                         save.sdrep = FALSE)
  
  saveRDS(mod, file.path(sub.dir,sprintf("Mod4_%03d.RDS",i)))
}

stopCluster(cluster)
```

### 14. EM5
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
  
  n_stocks = n_regions = n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model=rep("equilibrium",n_stocks),
                  sigma=rep("rec+1",n_stocks),
                  cor=rep("iid",n_stocks),
                  NAA_where = om$input$data$NAA_where)
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  mod <- loop_through_fn(om = om_with_data, 
                         em_info = info,
                         random = random,
                         M_em = M_em, 
                         sel_em = sel_em,
                         NAA_re_em = NAA_re_em,
                         move_em = move,
                         age_comp_em = "multinomial",
                         em.opt = list(separate.em = FALSE, separate.em.type = NULL, 
                                       do.move = TRUE, est.move = FALSE),
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 20, 
                         hcr.type = 1,
                         seed = 123+i,
                         save.sdrep = FALSE)
  
  saveRDS(mod, file.path(sub.dir,sprintf("Mod5_%03d.RDS",i)))
}

stopCluster(cluster)
```

### 15. EM6
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i, random = random)
  
  n_stocks = n_regions = n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model=rep("equilibrium",n_stocks),
                  sigma=rep("rec+1",n_stocks),
                  cor=rep("iid",n_stocks),
                  NAA_where = om$input$data$NAA_where)
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  mod <- loop_through_fn(om = om_with_data, 
                         em_info = info,
                         random = random,
                         M_em = M_em, 
                         sel_em = sel_em,
                         NAA_re_em = NAA_re_em,
                         move_em = move,
                         age_comp_em = "multinomial",
                         em.opt = list(separate.em = FALSE, separate.em.type = NULL, 
                                       do.move = TRUE, est.move = TRUE),
                         assess_years = assess.years, 
                         assess_interval = assess.interval, 
                         base_years = base.years,
                         year.use = 20, 
                         hcr.type = 1,
                         seed = 123+i,
                         save.sdrep = FALSE,
                         save.last.em = TRUE)
  
  saveRDS(mod, file.path(sub.dir,sprintf("Mod6_%03d.RDS",i)))
}

stopCluster(cluster)
```

### 16. Collect simulation results
```{r, eval=FALSE}
mods = list()

for (nsim in 1:10) {
  file.names <- file.path(sub.dir,paste0("Mod",1:6,sprintf("_%03d.RDS",nsim)))
  mods[[nsim]] <- lapply(file.names, readRDS)
}
# Please save the results for Ex6:6.Performance-Analysis-Tools

plot_mse_output(mods, base.mod = 6, out.type = "pdf") # base.mod = correct EM
# plot_mse_output(mods,  base.mod = 6, out.type = "png") # Alternative
# plot_mse_output(mods,  base.mod = 6, out.type = "html") # Alternative
```

