---
title: "Parallel Computing"
author: "Chengxue Li"
date: "Last update: `r Sys.Date()`"
output: 
  html_document:
    includes:
      in_header: header.html
editor_options: 
  markdown: 
    wrap: 72
---

This vignette includes an example of how to use parallel computing to conduct management strategy evaluation.

### 1. Load "WHAM" and "whamMSE":

```{r, eval=FALSE}
library(wham)
library(whamMSE)

main.dir = here::here()
```

### 2. Generate basic information

```{r, eval=FALSE}
year_start  <- 2003  # starting year in the burn-in period
year_end    <- 2022  # end year in the burn-in period
MSE_years   <- 3     # number of years in the feedback loop
# Note: no need to include MSE_years in simulation-estimation 

info <- generate_basic_info(n_stocks   = 2, 
                            n_regions  = 2, 
                            n_indices  = 2, 
                            n_fleets   = 2, 
                            n_seasons  = 4, 
                            base.years = year_start:year_end, 
                            n_feedback_years = MSE_years, 
                            life_history  = "medium", 
                            n_ages        = 12, 
                            Fbar_ages     = 12, 
                            recruit_model = 2, 
                            F_info     = list(F.year1 = 0.2, Fhist = "updown",Fmax = 1, Fmin = 0.2,change_time = 0.5),
                            catch_info = list(catch_cv = 0.1, catch_Neff = 200), 
                            index_info = list(index_cv = 0.1, index_Neff = 200, fracyr_indices = 0.5, q = 0.2), 
                            fracyr_spawn = 0.5, 
                            bias.correct.process     = FALSE, 
                            bias.correct.observation = FALSE, 
                            bias.correct.BRPs        = FALSE, 
                            mig_type = 0) 

basic_info = info$basic_info # collect basic information
catch_info = info$catch_info # collect fleet catch information
index_info = info$index_info # collect survey information
F_info = info$F # collect fishing information

# see more details using ?generate_basic_info
```

### 3. Specify movement type and movement rate

```{r, eval=FALSE}
basic_info <- generate_NAA_where(basic_info = basic_info, move.type = 2) # "bidirectional" movement

move <- generate_move(basic_info = basic_info, move.type = 2, move.rate = 0.3, move.re = "constant")

```

### 4. Configure selecitvity and natural mortality

```{r, eval=FALSE}
n_stocks  <- as.integer(basic_info['n_stocks'])
n_regions <- as.integer(basic_info['n_regions'])
n_fleets  <- as.integer(basic_info['n_fleets'])
n_indices <- as.integer(basic_info['n_indices'])
n_ages    <- as.integer(basic_info['n_ages'])

# Selectivity Configuration
fleet_pars <- c(5,1)
index_pars <- c(2,1)
sel <- list(model=rep("logistic",n_fleets+n_indices),
            initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

# M Configuration
M <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
```

### 5. Configure numbers-at-age (NAA)

```{r, eval=FALSE}
sigma        <- "rec+1"
re_cor       <- "iid"
ini.opt      <- "equilibrium" # option <- c("age-specific-fe", "equilibrium")
Rec_sig      <- 0.5 # (sigma for recruitment)
NAA_sig      <- 0.5 # (sigma for NAA)

sigma_vals = array(NAA_sig, dim = c(n_stocks, n_regions, n_ages)) # n_stocks x n_regions x n_ages
sigma_vals[,,1] = Rec_sig

# Set initial NAA for each stock
log_N1    <- rep(10,n_stocks)
log_N1[1] <- log(exp(10)*2) # N1_stock1 is 2 times higher than N1_stock2
N1_pars   <- generate_ini_N1(log_N1,basic_info,ini.opt)

# Set mean recruitment para. for each stock
mean_rec_par <- list()
for (i in 1:n_stocks) mean_rec_par[[i]] <- exp(log_N1[i])

NAA_re <- list(N1_model=rep(ini.opt,n_stocks),
               sigma=rep(sigma,n_stocks),
               cor=rep(re_cor,n_stocks),
               recruit_model = 2,  # rec random around the mean
               recruit_pars = mean_rec_par, 
               sigma_vals = sigma_vals,  
               N1_pars = N1_pars,
               NAA_where = basic_info$NAA_where) # NAA_where must be specified in basic_info!
```

### 6. Generate wham input

Here we use `prepare_wham_input()` function to generate a wham input
using the basic information we set above:

```{r, eval=FALSE}
input <- prepare_wham_input(basic_info = basic_info, selectivity = sel, M = M, NAA_re = NAA_re, move = move, catch_info = catch_info, index_info = index_info, F = F_info)
```

### 7. Generate operating model

```{r, eval=FALSE}
random = input$random # check what processes are random effects
input$random = NULL # so inner optimization won't change simulated RE
om <- fit_wham(input, do.fit = F, do.brps = T, MakeADFun.silent = TRUE)
# Note: do.fit must be FALSE (no modeling fitting yet)
```

### 8. Specify assessment interval and assessment year in the feedback loop

```{r, eval=FALSE}
assess.interval <- 3 # Note: assessment interval is 3 years, given the feedback period is 3 years, there will be only 1 assessment
base.years      <- year_start:year_end # Burn-in period
first.year      <- head(base.years,1)
terminal.year   <- tail(base.years,1)
assess.years    <- seq(terminal.year, tail(om$years,1)-assess.interval,by = assess.interval)

mods <- list() # Create a list to save MSE outputs
```

### 9. Parallel computing (EM1)

```{r, eval=FALSE}
library(doParallel)
library(foreach)
detectCores() # check how many cores available
cluster <- makeCluster(10) 
registerDoParallel(cluster)

results <- list()

results <- foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i)
  
  n_stocks = n_regions = n_fleets = n_indices = 1
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
                 initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  tryCatch({
    mod <- loop_through_fn(om = om_with_data, 
                            random = random,
                            M_om = M,
                            sel_om = sel, 
                            NAA_re_om = NAA_re, 
                            move_om = move,
                            age_comp_om = "multinomial",
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial", # Note: not multinomial!
                            em.opt = list(separate.em = TRUE, separate.em.type = 1, do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, 
                            seed = 123+i,
                            save.sdrep = FALSE)
    results[[i]] <- mod
    
  }, error = function(e) {
    
    warning(paste("Iteration",i,"failed","e$message"))
    
    results[[i]] <- NULL
    
  })
  
}
mods[[1]] <- results

stopCluster(cluster)
```

### 10. Parallel computing (EM2)
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

results <- list()

results <- foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i)
  
  n_stocks = n_regions = 1
  n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")

  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  tryCatch({
    mod <- loop_through_fn(om = om_with_data, 
                            random = random,
                            M_om = M,
                            sel_om = sel, 
                            NAA_re_om = NAA_re, 
                            move_om = move,
                            age_comp_om = "multinomial",
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 2, do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            seed = 123+i,
                            save.sdrep = FALSE)
  results[[i]] <- mod
    
  }, error = function(e) {
    
    warning(paste("Iteration",i,"failed","e$message"))
    
    results[[i]] <- NULL
    
  })
  
}
mods[[2]] <- results

stopCluster(cluster)
```

### 10. Parallel computing (EM3)
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

results <- list()

results <- foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i)
  
  n_stocks = n_regions = n_fleets = n_indices = 1
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model="equilibrium",sigma="rec+1",cor="iid")
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  tryCatch({
    mod <- loop_through_fn(om = om_with_data, 
                            random = random,
                            M_om = M,
                            sel_om = sel, 
                            NAA_re_om = NAA_re, 
                            move_om = move,
                            age_comp_om = "multinomial",
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = TRUE, separate.em.type = 3, do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            seed = 123+i,
                            save.sdrep = FALSE)
  results[[i]] <- mod
    
  }, error = function(e) {
    
    warning(paste("Iteration",i,"failed","e$message"))
    
    results[[i]] <- NULL
    
  })
  
}
mods[[3]] <- results

stopCluster(cluster)
```

### 10. Parallel computing (EM4)
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

results <- list()

results <- foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i)
  
  n_stocks = n_regions = n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model=rep("equilibrium",n_stocks),
                  sigma=rep("rec+1",n_stocks),
                  cor=rep("iid",n_stocks))
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  tryCatch({
    mod <- loop_through_fn(om = om_with_data, 
                            random = random,
                            M_om = M,
                            sel_om = sel, 
                            NAA_re_om = NAA_re, 
                            move_om = move,
                            age_comp_om = "multinomial",
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = move,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = FALSE, separate.em.type = 3, do.move = TRUE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            seed = 123+i,
                            save.sdrep = FALSE)
  results[[i]] <- mod
    
  }, error = function(e) {
    
    warning(paste("Iteration",i,"failed","e$message"))
    
    results[[i]] <- NULL
    
  })
  
}
mods[[4]] <- results

stopCluster(cluster)
```

### 10. Parallel computing (EM5)
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

results <- list()

results <- foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i)
  
  n_stocks = n_regions = n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))

  NAA_re_em <- list(N1_model=rep("equilibrium",n_stocks),
                  sigma=rep("rec+1",n_stocks),
                  cor=rep("iid",n_stocks),
                  recruit_model = 2)

  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  tryCatch({
    mod <- loop_through_fn(om = om_with_data, 
                            random = random,
                            M_om = M,
                            sel_om = sel, 
                            NAA_re_om = NAA_re, 
                            move_om = move,
                            age_comp_om = "multinomial",
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = FALSE, separate.em.type = 3, do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            seed = 123+i,
                            save.sdrep = FALSE)
  results[[i]] <- mod
    
  }, error = function(e) {
    
    warning(paste("Iteration",i,"failed","e$message"))
    
    results[[i]] <- NULL
    
  })
  
}
mods[[5]] <- results

stopCluster(cluster)
```

### 10. Parallel computing (EM6)
```{r, eval=FALSE}
cluster <- makeCluster(10) 
registerDoParallel(cluster)

results <- list()

results <- foreach (i = 1:10) %dopar% {
  
  library(wham)
  library(whamMSE)
  
  om_with_data <- update_om_fn(om, seed = 123+i)
  
  n_stocks = n_regions = n_fleets = n_indices = 2
  
  sel_em <- list(model=rep("logistic",n_fleets+n_indices),
               initial_pars=c(rep(list(fleet_pars),n_fleets),rep(list(index_pars),n_indices)))
  
  NAA_re_em <- list(N1_model=rep("equilibrium",n_stocks),
                  sigma=rep("rec",n_stocks),
                  cor=rep("iid",n_stocks))
  
  M_em <- list(model="constant",initial_means=array(0.2, dim = c(n_stocks,n_regions,n_ages)))
  
  tryCatch({
    mod <- loop_through_fn(om = om_with_data, 
                            random = random,
                            M_om = M,
                            sel_om = sel, 
                            NAA_re_om = NAA_re, 
                            move_om = move,
                            age_comp_om = "multinomial",
                            M_em = M_em, 
                            sel_em = sel_em, 
                            NAA_re_em = NAA_re_em, 
                            move_em = NULL,
                            age_comp_em = "multinomial",
                            em.opt = list(separate.em = FALSE, separate.em.type = 3, do.move = FALSE, est.move = FALSE),
                            assess_years = assess.years, 
                            assess_interval = assess.interval, 
                            base_years = base.years,
                            year.use = 20, # number of years of data you want to use in the assessment model
                            seed = 123+i,
                            save.sdrep = FALSE)
  results[[i]] <- mod
    
  }, error = function(e) {
    
    warning(paste("Iteration",i,"failed","e$message"))
    
    results[[i]] <- NULL
    
  })
  
}
mods[[6]] <- results

stopCluster(cluster)
```