---
title: "Comprehensive-Output-Analysis-Tools"
author: "Chengxue Li"
date: "Last update: `r Sys.Date()`"
output: 
  html_document:
    includes:
      in_header: header.html
editor_options: 
  markdown: 
    wrap: 72
---

This vignette introduces how to use `plot_mse_output()` function to
summarize and visualize the results of management strategy evaluation.
`plot_mse_output()` function is capable of summarizing management
strategy evaluation results from one realization and from n
realizations, when using parallel computing (see Example 5 for more
details)). The results can be presented in line plot, scatter plot, box
plot, and radar chart for users to better compare performance among
different management strategies. Please use ?plot_mse_output to see more
details. Below are one example about plotting results from one
realization and one example about plotting results from n realizations.

### 1. Results from one realization

Suppose you only run a management strategy evaluation for one
realization (e.g. Example 4):

```{r, eval=FALSE}
library(wham)
library(whamMSE)

main.dir = here::here() # Or the path you want to save the results.

# Remember in Example 4 we have 2 EMs saved in a list called "mods". 

plot_mse_output(mods, use.n.years = 10) # also summarize the performance from the last 10 model years

```

![Figure 1](Report(1)/SSB.PNG) ![Figure 2](Report(1)/Fbar.PNG) ![Figure
3](Report(1)/Catch_r.PNG) ![Figure 4](Report(1)/Catch_s.PNG) ![Figure
5](Report(1)/Performance_last_10years.PNG)

### 2. Results from n realizations

Suppose you run management strategy evaluation for n realizations
following Example 5, you can use the code below to plot the results.
Remember mods[[1]] is a list that contains results of the performance of
EM1 in different realizations.

```{r, eval=FALSE}
mods_list <- list()

for (i in 1:length(mods[[1]])) { # check the number of realizations you have
  mods_list[[i]] <- list(mods[[1]][[i]],mods[[2]][[i]],mods[[3]][[i]],
                         mods[[4]][[i]],mods[[5]][[i]],mods[[6]][[i]]) # collect performance from EMs in every realization 
}

plot_mse_output(mods = mods_list, 
                main.dir = getwd(), 
                use.n.years = 10, # The last 10 years will be used for comparison of long-term performance 
                base.mod = 4, # "correct" EM that can be used as a reference
                short.term = NULL, # if NULL, results of first 2 years will be used for comparison of short-term performance
                dpi = 150,
                out.type = "pdf") # Options: png, pdf, html 
```

![Figure 1](Report(2)/Performance_last_10years.PNG)

![Figure 2](Report(2)/Performance_Relative_Diff.png)

![Figure 3](Report(2)/Radar_Chart.png)

![Figure 4](Report(2)/Population_Status_all.png)

![Figure 5](Report(2)/Performance_First_N_Years.png)

![Figure 6](Report(2)/Simulation_Estimation.png)

![Figure 7](Report(2)/AIC.png)
