---
title: "Comprehensive Output Analysis Tools"
author: "Chengxue Li"
date: "Last update: `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    includes:
      in_header: header.html
editor_options: 
  markdown: 
    wrap: 72
---

This vignette introduces how to use `plot_mse_output()` function to
summarize and visualize the results of management strategy evaluation.
`plot_mse_output()` function is capable of summarizing management
strategy evaluation results from one realization and from n
realizations, when using parallel computing (see Example 5 for more
details)). The results can be presented in line plot, scatter plot, box
plot, and radar chart for users to better compare performance among
different management strategies. Please use ?plot_mse_output to see more
details. Below are one example about plotting results from one
realization and one example about plotting results from n realizations.

### 1. Results from one realization

Suppose you only run a management strategy evaluation for one
realization (e.g. Example 4):

```{r, eval=FALSE}
library(wham)
library(whamMSE)

main.dir = here::here() # Or the path you want to save the results.

# Remember in Example 4 we have 2 EMs saved in a list called "mods". 

plot_mse_output(mods, use.n.years = 10) # also summarize the performance from the last 10 model years

```

#### 1.1 Spawning stock biomass (SSB) over time, stocks-specific SSB and global SSB are plotted below:
![Figure 1](Report(1)/SSB.PNG) 

#### 1.2 Fishing mortality (F) over time, region-specific Fmax and total F (sum of region-specific F) are plotted below:
![Figure 2](Report(1)/Fbar.PNG) 

#### 1.3 Catch over time, region-specific catch and total catch are plotted below:
![Figure 3](Report(1)/Catch_r.PNG) 

#### 1.4 Catch over time, stock-specific catch and total catch are plotted below:
![Figure 4](Report(1)/Catch_s.PNG) 

#### 1.5 Performance summarized over the last 10 years are summarized below:
![Figure 5](Report(1)/Performance_last_10years.PNG)

### 2. Results from n realizations

Suppose you have run management strategy evaluation for n realizations
(see Example 5 for more details), you can use the code below to plot the results.
Remember mods[[1]] is a list that contains results of the performance of
EM1 in different realizations.

```{r, eval=FALSE}
mods_list <- list()

for (i in 1:length(mods[[1]])) { # check the number of realizations you have
  mods_list[[i]] <- list(mods[[1]][[i]],mods[[2]][[i]],mods[[3]][[i]],
                         mods[[4]][[i]],mods[[5]][[i]],mods[[6]][[i]]) # collect performance from EMs in every realization 
}

plot_mse_output(mods = mods_list, 
                main.dir = getwd(), 
                use.n.years = 10, # The last 10 years will be used for comparison of long-term performance 
                base.mod = 4, # "correct" EM that can be used as a reference
                short.term = NULL, # if NULL, results of first 2 years will be used for comparison of short-term performance
                dpi = 150,
                out.type = "pdf") # Options: png, pdf, html 
```

#### 2.1 Performance summarized over the last 10 years across EMs over n realizations:
![Figure 1](Report(2)/Performance_last_10years.PNG)

#### 2.2 Relative difference in performance between correct EMs and mis-specified EMs:
![Figure 2](Report(2)/Performance_Relative_Diff.png)

#### 2.3 Holistic view of performance of different EMs:
![Figure 3](Report(2)/Radar_Chart.png)

#### 2.4 Annual population status in the last 10 years across EMs and n realizations:
![Figure 4](Report(2)/Population_Status_all.png)

#### 2.5 Short-term performance (over the first 2 years) across EMs and n realizations:
![Figure 5](Report(2)/Performance_First_N_Years.png)

#### 2.6 Estimates of mean recruitment and variances of recruitment and survival deviations:
![Figure 6](Report(2)/Simulation_Estimation.png)

#### 2.7 Margial AIC calculated for each EM:
![Figure 7](Report(2)/AIC.png)

#### 2.8 For full report of terminal-year assessment model, please use:
```{r, eval=FALSE}
plot_wham_output(mods_list[[1]][[1]], out.type = "html") # plot EM1 from the first realization
```

#### 2.9 Check model convergence in every assessment model in the feedback period
```{r, eval=FALSE}
mods_list[[1]][[2]]$converge_list # EM2 from the first realization
# 2 indicates the assessment model is converged.
```

#### 2.10 Check catch advice generated from every assessment model in the feedback period
```{r, eval=FALSE}
mods_list[[1]][[4]]$catch_advice # catch advice generated from EM4 from the first realization
```
